Practical Security Log Book

### 2014-12-15 logbook.tex
Written first lab report.

\title{The Folly of Passwords}
\author{Ross Fenning}

@@

\chapter{Wireshark Analysis}

@@

\noident and the ``Hello'' message is displayed. Note that you have to
provide your passphrase, which means your private key is involved in
decrypting the message. A malicious party would have to obtain both
your private key and the passphrase to use it in order to read this
encrypted message.

@@

>>>>

\title{Practical Security Log Book}
\author{Ross Fenning}

@@

\chapter{Wireshark Analysis of a Teardrop attack}

In this report, we look at a very old attack at the IP layer
known as the \emph{Teardrop} attack which was able to crash
Linux and Windows computers at the time of discovery.
We also look at the protocol violation used to achieve this attack
in the \emph{Wireshark} analysis tool.

\section{What is a Teardrop attack?}

The Teardrop attack \parencite{center1997cert} was discovered
in late 1997 as a potential vulnerability in Windows 3.1x, Windows 95 and Windows NT
systems and versions of Linux prior to versions 2.0.32 and 2.1.63.
The vulnerability could allow a remote attacker to crash computers
running these systems due to a bug in the code performing IP fragmentation
reassembly.

IP fragmentation \parencite{postel1981rfc,fall2011tcp} is necessary if a router
receives a packet larger than the maximum transmission unit (MTU) of the
next hop. When this occurs, a router has the option to break the datagram
into smaller fragments. RFC 815 \parencite{clark1982ip} describes
some algorithms for how downstream hops can then reassemble the fragmented
IP packets.

The Teardrop attack involves sending fragmented IP packets such that the offset
of later packets do not correspond with the length of earlier packets. Normally,
if the first fragment is 16 bytes in size, then the second fragment should
indicate an offset of 16 bytes. Setting the second packet to a lower
offset such that the packets seem to overlap was known to cause bugs in
these old operating systems and thus they became vulnerable to the Teardrop
attack.

\section{Example}

Appendix~\ref{appendix:packets} shows raw exports from Wireshark of two frames
found in a capture of a Teardrop attack. We can see that Wireshark has identified
frames 8 and 9 as being two IP fragments which together make up a single UDP
packet.

The IPv4 header advertises a total length for the first fragment (frame 8)
of 56 bytes. The frame comprises 20 bytes for the IPv4 header and 36 bytes of
data. The bytes \texttt{00 38} at offset \texttt{0010} show this header length.
The third bit is set on the bytes \texttt{20 00}, which indicates more
fragments are to come, but this fragment starts at offset 0 (i.e. it's the first
fragment). This is a fairly standard initial fragment and there is nothing
malformed in this frame.

The attack comes in the next fragment (frame 9) where bytes \texttt{00 03}
claim that this fragment has an offset of 3 eight-byte blocks, or 24 bytes.
This is not in accord with the fact that the previous fragment was 36 bytes
long and looks as if this fragment overlaps the previous one.

Wireshark
has reassembled this to an IPv4 packet with 28 bytes of data as it has
resolved the malformed fragmentation by losing the last 12 bytes of the
first fragment (i.e. scrolling back to the advertised offset of 24 and
then adding on the 4 bytes contained in the second fragment). It has
also identified that the resulting 28 byte payload length does not
agree with the 36 byte header claimed in the UDP header (bytes \texttt{00 24}
at offset \texttt{0026}) and flagged that as an error.

The older versions of Windows and Linux were vulnerable to bugs in
their fragmentation reassembly that would simply cause the operating
systems to crash when this overlap occurred.

@@

\noindent and the ``Hello'' message is displayed. Note that you have to
provide your passphrase, which means your private key is involved in
decrypting the message. A malicious party would have to obtain both
your private key and the passphrase to use it in order to read this
encrypted message.

@@

\appendix

\chapter{Teardrop Attack Packets}
\label{appendix:packets}

\lstinputlisting[basicstyle=\scriptsize]{frame8.txt}
\pagebreak
\lstinputlisting[basicstyle=\scriptsize]{frame9.txt}
